---
title: Evaluation of property price fluctuaction according to geographical landmarks, an Dublin case study
author:
  - name: Damien Dupré
    email: damien.dupre@dcu.ie
    affiliation: Dublin City University
    footnote: Corresponding Author
address:
  - code: Dublin City University
    address: Business School, Glasnevin, Dublin 9, Ireland 
abstract: |
  This is the abstract.

  It consists of two paragraphs.

journal: "An awesome journal"
date: "`r Sys.Date()`"
bibliography: mybibfile.bib
output: rticles::elsevier_article
header-includes:
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{array}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  cache.extra = knitr::rand_seed, 
  message = FALSE, 
  warning = FALSE, 
  error = FALSE, 
  echo = FALSE, 
  fig.align="center", 
  fig.pos = 'H')
# options
options(scipen=999)
# tools
library(here)
library(lubridate)
library(readr)
library(papaja)
# data wrangling
library(plyr)
library(dplyr)
# table design
library(kableExtra)
# figure design
library(ggplot2)
library(ggpubr)
library(ggridges)
# map tools
library(sf)
library(tmap)
library(lwgeom)
library(osmdata)
library(OpenStreetMap)
# statistical analysis
library(mgcv)
library(xgboost)
# custom function
source(here::here("script/my_viz.R"))
```


Introdution
==========================

To acquire a property is one of the most important achievement that individuals are seeking. It provides not only a housing security but also the feeling of being a landowner. However the access to the status of landowner is complicated because buying a property is the most expensive spending of in a lifetime. For this reason understanding the factors which are explaining how property prices evolve is a necessity.


```{r all_OSM_features}
read_csv(here::here("data/OSM_dublin_features.csv")) %>% 
  kable(
    "latex",
    caption = "Relevant Open Street Map features for dublin area.",
    booktabs = T,
    linesep = ""
  )  %>% 
  kable_styling(font_size = 8,
                full_width = FALSE) %>%
  column_spec(2,width = "3in")
```
